---
title: "Wytham Rodent Faecal Virome"
author: "Jayna"
date: "25/11/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, warning = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(vegan)
library(ggvenn)
library(patchwork)
source("read_and_process_data.R")
source("diversityStats.R")

MNKA <- read_csv("../rodent_borne_viruses/wytham_woods_rodents/MNKA_wytham_since_Nov2016.csv")

```

## Raw viral abundance

```{r echo=FALSE, warning = FALSE, fig.align='center', fig.height = 1, fig.width=3, message=FALSE, warning=FALSE}

## Raw viral abundance 
viral_abundance <-
  total_virus_reads %>%
  mutate(Interval = as.character(Interval)) %>%
  left_join(samples_per_pool) %>%
  mutate(Species = fct_relevel(Species, c("AS", "AF", "MG")))

# viral_abundance$Species <- factor(viral_abundance$Species, levels=c("AS", "AF", "MG"))


options(scipen = 999)

raw_abundance.p <- 
  viral_abundance %>%
  ggplot(aes(dmy(collection_date), `Proportion of viral reads`, 
             colour=Species, fill=Species, group=Species))+
  geom_point()+geom_line()+
  theme_bw()+
  xlab("Time period") +
  scale_color_manual(values=c("#66C2A5", "#8DA0CB", "#FC8D62")) +
  facet_grid(~Species)+
  geom_text(aes(label=`Number of individuals`, y = (`Proportion of viral reads`) + 0.001),
            position = position_dodge(0.9),
            vjust = -0.9, colour="black")+
  scale_y_continuous(limits=c(0,0.07))+
  scale_x_date(limits =c(dmy("20/01/2017"), dmy("15/12/2017")))+
  ylab("Proportion of viral reads") + xlab("Time")+
  theme(legend.position = "none")

raw_abundance.p

# ggsave("../Wytham_rodent_manuscript/revisions/new/Raw_abundance_2022-06-10.pdf", height = 2.5, width = 6.5, plot = raw_abundance.p)

# Statistics

max <- max(total_virus_reads$Sequences)
min <-  min(total_virus_reads$Sequences)

print(paste0("max reads per pool ", max))
print(paste0("min reads per pool ", min))

AS_mean <- median(subset(total_virus_reads, Species=="AS")$Sequences)
AF_mean <- median(subset(total_virus_reads, Species=="AF")$Sequences)
MG_mean <- median(subset(total_virus_reads, Species=="MG")$Sequences)

print(paste0("median reads per AS pool ", AS_mean))
print(paste0("median reads per AF pool ", AF_mean))
print(paste0("median reads per MG pool ", MG_mean))

# proportion of viral reads (denominator total raw reads i.e. paired-end x 2)
prop_viral_reads <- (sum(total_virus_reads$Sequences)/(2*355917017))*100

sum(total_virus_reads$Sequences)

print(paste0("Percentage viral reads ", round(prop_viral_reads,digits = 4), "%"))

length(unique(all_virus.df$virus_family))

all.hist <-
  all_virus_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000) %>%
  group_by(virus_species, Species) %>%
  summarise(count = n()) %>%
  ggplot(aes(count)) + geom_histogram(fill="grey", colour="black", bins = 5, aes(y = stat(count) / sum(count)))  + theme_bw() + xlab("Number of intervals detected") + ylab("Proportion")  + ggtitle("All") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.5))

vert.hist <-
  verterb_virus_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000) %>%
  group_by(virus_species, Species) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=count)) + geom_histogram(fill="grey", colour="black", bins = 5, aes(y = stat(count) / sum(count))) + xlab("Number of intervals detected") + ylab("Proportion") + theme_bw()  + ggtitle("Vertebrate-associated") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.5))

phage.hist <-
  phage_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000) %>%
  group_by(virus_species, Species) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=count)) + geom_histogram(fill="grey", colour="black", bins = 5, aes(y = stat(count) / sum(count))) + xlab("Number of intervals detected") + ylab("Proportion") + theme_bw() + ggtitle("Bacteriophage") +  theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.5)) 

prop_vs_n_intervals <- ggarrange(all.hist, vert.hist, phage.hist, nrow =1)
prop_vs_n_intervals


all.interval.hist <-
  all_virus_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000) %>%
  group_by(Interval) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=Interval)) + geom_col(fill="grey", colour="black", aes(y = (count)/sum(count)))  + theme_bw() + xlab("Sampling period") + ylab("Proportion")  + ggtitle("All") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.4))

vert.interval.hist <-
  verterb_virus_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000) %>%
  group_by(Interval) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=Interval)) + geom_col(fill="grey", colour="black", aes(y = (count)/sum(count)))  + theme_bw() + xlab("Sampling period") + ylab("Proportion")  + ggtitle("Vertebrate-associated") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.4))

phage.interval.hist <-
  phage_species.df %>%
  filter(viral_read_count/(`Number of paired-end reads`*2) >= 1/10000000)%>%
  group_by(Interval) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=Interval)) + geom_col(fill="grey", colour="black", aes(y = (count)/sum(count)))  + theme_bw() + xlab("Sampling period") + ylab("Proportion")  + ggtitle("Bacteriophage") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,0.4))

prop_virus_vs_sampling_period <-
  ggarrange(all.interval.hist, vert.interval.hist, phage.interval.hist, nrow=1)
prop_virus_vs_sampling_period
combined_summary <- ggarrange(prop_vs_n_intervals, prop_virus_vs_sampling_period, nrow=2)
combined_summary

# ggsave("../Wytham_rodent_manuscript/revisions/new/prop_virus_detected_vs_number_of_intervals.pdf", combined_summary, height=5, width=8)

ggarrange(prop_vs_n_intervals, prop_virus_vs_sampling_period, nrow=2)


phage_species.df %>%
  filter(viral_read_count/((`Number of paired-end reads`)*2) >= 1/10000000)%>%
  group_by(virus_species, Species) %>%
  summarise(count = n()) %>%
  group_by(count) %>%
  summarise(count2 = n())


all_virus_species.df %>%
  filter(viral_read_count/((`Number of paired-end reads`)*2) >= 1/10000000)%>%
  group_by(Interval) %>%
  summarise(count = n()) %>%
  filter(Interval >= 1)%>%
  summarise(sum = sum(count))


```

## Filtered contingency table (mapped read counts)

```{r}

# filtering the contingency table for contigs where abundance is greater than 1 per 10 million read threshold

contingency_table.s <-
  annotate_and_filter_contingency_table(contingency_table, blast_results_table) %>%
  
  mutate(AS1_abundance =
           round(WTCHG_714232_72815161_AS_1 / ((subset(
             samples_per_pool, Species == "AS" &
               Interval == 1
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AS2_abundance =
           round(WTCHG_714232_72835163_AS_2 / ((subset(
             samples_per_pool, Species == "AS" &
               Interval == 2
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AS3_abundance =
           round(WTCHG_714232_72865166_AS_3 / ((subset(
             samples_per_pool, Species == "AS" &
               Interval == 3
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AS4_abundance =
           round(WTCHG_714232_71935169_AS_4 / ((subset(
             samples_per_pool, Species == "AS" &
               Interval == 4
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AS5_abundance =
           round(WTCHG_714232_71965172_AS_5 / ((subset(
             samples_per_pool, Species == "AS" &
               Interval == 5
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(MG1_abundance =
           round(WTCHG_714232_72825162_MG_1 / ((subset(
             samples_per_pool, Species == "MG" &
               Interval == 1
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(MG2_abundance =
           round(WTCHG_714232_72845164_MG_2 / ((subset(
             samples_per_pool, Species == "MG" &
               Interval == 2
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(MG3_abundance =
           round(WTCHG_714232_72875167_MG_3 / ((subset(
             samples_per_pool, Species == "MG" &
               Interval == 3
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(MG4_abundance =
           round(WTCHG_714232_71945170_MG_4 / ((subset(
             samples_per_pool, Species == "MG" &
               Interval == 4
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(MG5_abundance =
           round(WTCHG_714232_71975173_MG_5 / ((subset(
             samples_per_pool, Species == "MG" &
               Interval == 5
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AF3_abundance =
           round(WTCHG_714232_72855165_AF_3 / ((subset(
             samples_per_pool, Species == "AF" &
               Interval == 3
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AF4_abundance =
           round(WTCHG_714232_72885168_AF_4 / ((subset(
             samples_per_pool, Species == "AF" &
               Interval == 4
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  mutate(AF5_abundance =
           round(WTCHG_714232_71955171_AF_5 / ((subset(
             samples_per_pool, Species == "AF" &
               Interval == 5
           ))$`Number of paired-end reads`*2), digits=8)) %>%
  
  mutate(WTCHG_714232_72815161_AS_1 =
           ifelse((AS1_abundance <= (1 / 10000000)), 0, WTCHG_714232_72815161_AS_1)) %>%
  mutate(WTCHG_714232_72835163_AS_2 =
           ifelse((AS2_abundance <= (1 / 10000000)), 0, WTCHG_714232_72835163_AS_2)) %>%
  mutate(WTCHG_714232_72865166_AS_3 =
           ifelse((AS3_abundance <= (1 / 10000000)), 0, WTCHG_714232_72865166_AS_3)) %>%
  mutate(WTCHG_714232_71935169_AS_4 =
           ifelse((AS4_abundance <= (1 / 10000000)), 0, WTCHG_714232_71935169_AS_4)) %>%
  mutate(WTCHG_714232_71965172_AS_5 =
           ifelse((AS5_abundance <= (1 / 10000000)), 0, WTCHG_714232_71965172_AS_5)) %>%
  mutate(WTCHG_714232_72825162_MG_1 =
           ifelse((MG1_abundance <= (1 / 10000000)), 0, WTCHG_714232_72825162_MG_1)) %>%
  mutate(WTCHG_714232_72845164_MG_2 =
           ifelse((MG2_abundance <= (1 / 10000000)), 0, WTCHG_714232_72845164_MG_2)) %>%
  mutate(WTCHG_714232_72875167_MG_3 =
           ifelse((MG3_abundance <= (1 / 10000000)), 0, WTCHG_714232_72875167_MG_3)) %>%
  mutate(WTCHG_714232_71945170_MG_4 =
           ifelse((MG4_abundance <= (1 / 10000000)), 0, WTCHG_714232_71945170_MG_4)) %>%
  mutate(WTCHG_714232_71975173_MG_5 =
           ifelse((MG5_abundance <= (1 / 10000000)), 0, WTCHG_714232_71975173_MG_5)) %>%
  mutate(WTCHG_714232_72855165_AF_3 =
           ifelse((AF3_abundance <= (1 / 10000000)), 0, WTCHG_714232_72855165_AF_3)) %>%
  mutate(WTCHG_714232_72885168_AF_4 =
           ifelse((AF4_abundance <= (1 / 10000000)), 0, WTCHG_714232_72885168_AF_4)) %>%
  mutate(WTCHG_714232_71955171_AF_5 =
           ifelse((AF5_abundance <= (1 / 10000000)), 0, WTCHG_714232_71955171_AF_5)) %>%
  
  filter(type != "NA")

```

# Venn diagrams / contigs / virus families

```{r, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}

# total_virus_reads <- total_virus_reads %>% rename(Sequences = `# Sequences`)
   

get_venn_data <- function(blast_results_table, contingency_table, list) {
  blast_contigs_subset <- blast_results_table %>%
    filter(`length` >= 200)
  
  
  contigency_table.s <-
    annotate_and_filter_contingency_table(contingency_table, blast_contigs_subset) %>%
    filter(virus_family %in% list) %>%
    
    rename(AS1 = WTCHG_714232_72815161_AS_1) %>%
    rename(AS2 = WTCHG_714232_72835163_AS_2) %>%
    rename(AS3 = WTCHG_714232_72865166_AS_3) %>%
    rename(AS4 = WTCHG_714232_71935169_AS_4) %>%
    rename(AS5 = WTCHG_714232_71965172_AS_5) %>%
    rename(MG1 = WTCHG_714232_72825162_MG_1) %>%
    rename(MG2 = WTCHG_714232_72845164_MG_2) %>%
    rename(MG3 = WTCHG_714232_72875167_MG_3) %>%
    rename(MG4 = WTCHG_714232_71945170_MG_4) %>%
    rename(MG5 = WTCHG_714232_71975173_MG_5) %>%
    rename(AF3 = WTCHG_714232_72855165_AF_3) %>%
    rename(AF4 = WTCHG_714232_72885168_AF_4) %>%
    rename(AF5 = WTCHG_714232_71955171_AF_5) %>%
    
    mutate(AS1_abundance =
             round(AS1 / ((subset(
               samples_per_pool, Species == "AS" &
                 Interval == 1
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AS2_abundance =
             round(AS2 / ((subset(
               samples_per_pool, Species == "AS" &
                 Interval == 2
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AS3_abundance =
             round(AS3 / ((subset(
               samples_per_pool, Species == "AS" &
                 Interval == 3
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AS4_abundance =
             round(AS4 / ((subset(
               samples_per_pool, Species == "AS" &
                 Interval == 4
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AS5_abundance =
             round(AS5 / ((subset(
               samples_per_pool, Species == "AS" &
                 Interval == 5
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(MG1_abundance =
             round(MG1 / ((subset(
               samples_per_pool, Species == "MG" &
                 Interval == 1
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(MG2_abundance =
             round(MG2 / ((subset(
               samples_per_pool, Species == "MG" &
                 Interval == 2
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(MG3_abundance =
             round(MG3 / ((subset(
               samples_per_pool, Species == "MG" &
                 Interval == 3
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(MG4_abundance =
             round(MG4 / ((subset(
               samples_per_pool, Species == "MG" &
                 Interval == 4
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(MG5_abundance =
             round(MG5 / ((subset(
               samples_per_pool, Species == "MG" &
                 Interval == 5
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AF3_abundance =
             round(AF3 / ((subset(
               samples_per_pool, Species == "AF" &
                 Interval == 3
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AF4_abundance =
             round(AF4 / ((subset(
               samples_per_pool, Species == "AF" &
                 Interval == 4
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    mutate(AF5_abundance =
             round(AF5 / ((subset(
               samples_per_pool, Species == "AF" &
                 Interval == 5
             ))$`Number of paired-end reads`*2), digits=8)) %>%
    
    mutate(AS1 =
             ifelse((AS1_abundance <= (1 / 10000000)), 0, AS1)) %>%
    mutate(AS2 =
             ifelse((AS2_abundance <= (1 / 10000000)), 0, AS2)) %>%
    mutate(AS3 =
             ifelse((AS3_abundance <= (1 / 10000000)), 0, AS3)) %>%
    mutate(AS4 =
             ifelse((AS4_abundance <= (1 / 10000000)), 0, AS4)) %>%
    mutate(AS5 =
             ifelse((AS5_abundance <= (1 / 10000000)), 0, AS5)) %>%
    mutate(MG1 =
             ifelse((MG1_abundance <= (1 / 10000000)), 0, MG1)) %>%
    mutate(MG2 =
             ifelse((MG2_abundance <= (1 / 10000000)), 0, MG2)) %>%
    mutate(MG3 =
             ifelse((MG3_abundance <= (1 / 10000000)), 0, MG3)) %>%
    mutate(MG4 =
             ifelse((MG4_abundance <= (1 / 10000000)), 0, MG4)) %>%
    mutate(MG5 =
             ifelse((MG5_abundance <= (1 / 10000000)), 0, MG5)) %>%
    mutate(AF3 =
             ifelse((AF3_abundance <= (1 / 10000000)), 0, AF3)) %>%
    mutate(AF4 =
             ifelse((AF4_abundance <= (1 / 10000000)), 0, AF4)) %>%
    mutate(AF5 =
             ifelse((AF5_abundance <= (1 / 10000000)), 0, AF5)) %>%
    
    rowwise(`contig ID`) %>%
    mutate(total_AS = sum(
      c(
        AS1,
        AS2,
        AS3,
        AS4,
        AS5
      )
    )) %>%
    mutate(total_MG = sum(
      c(
        MG1,
        MG2,
        MG3,
        MG4,
        MG5
      )
    )) %>%
    mutate(total_AF = sum(
      c(
        AF3,
        AF4,
        AF5
      )
    )) %>%
    filter(sum(total_AS, total_AF, total_MG) >= 20) %>%
    filter(type != "NA")
  
  
  AS <- list(
    AS1 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AS1)),
    AS2 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AS2)),
    AS3 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AS3)),
    AS4 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AS4)),
    AS5 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AS5))
  )
  
  MG <- list(
    MG1 = rep(contigency_table.s$`contig ID`, (contigency_table.s$MG1)),
    MG2 = rep(contigency_table.s$`contig ID`, (contigency_table.s$MG2)),
    MG3 = rep(contigency_table.s$`contig ID`, (contigency_table.s$MG3)),
    MG4 = rep(contigency_table.s$`contig ID`, (contigency_table.s$MG4)),
    MG5 = rep(contigency_table.s$`contig ID`, (contigency_table.s$MG5))
  )
  
  AF <- list(
    AF3 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AF3)),
    AF4 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AF4)),
    AF5 = rep(contigency_table.s$`contig ID`, (contigency_table.s$AF5))
  )
  
  return(list(AS = AS, AF = AF, MG =  MG))
}

get_viruscontigs_venn <- function() {
  
  vert_viruses <-
    get_venn_data(blast_results_table,
                  contingency_table,
                  unique(verteb_viruses.df$virus_family))
  invert_viruses <-
    get_venn_data(blast_results_table,
                  contingency_table,
                  unique(invert_virus.df$virus_family))
  plant_viruses <-
    get_venn_data(blast_results_table,
                  contingency_table,
                  unique(plant_virus.df$virus_family))
  bacterio_viruses <-
    get_venn_data(blast_results_table,
                  contingency_table,
                  unique(bacteriophage.df$virus_family))
  
  vert.venn <-
    list(
      AS = unlist(vert_viruses[[1]]),
      AF = unlist(vert_viruses[[2]]),
      MG = unlist(vert_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color = c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  invert.venn <-
    list(
      AS = unlist(invert_viruses[[1]]),
      AF = unlist(invert_viruses[[2]]),
      MG = unlist(invert_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color = c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  plant.venn <-
    list(
      AS = unlist(plant_viruses[[1]]),
      AF = unlist(plant_viruses[[2]]),
      MG = unlist(plant_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color = c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  bacterio.venn <-
    list(
      AS = unlist(bacterio_viruses[[1]]),
      AF = unlist(bacterio_viruses[[2]]),
      MG = unlist(bacterio_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color = c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  contigs.venn <-
    (vert.venn | invert.venn) / (plant.venn | bacterio.venn)
  
  return(contigs.venn)
}

contigs.venn <- get_viruscontigs_venn()

contigs.venn

all_virus <-
  get_venn_data(blast_results_table,
                contingency_table,
                unique(all_virus.df$virus_family))

all.venn <-
  list(
    AS = unlist(all_virus[[1]]),
    AF = unlist(all_virus[[2]]),
    MG = unlist(all_virus[[3]])
  ) %>%
  ggvenn(
    fill_color = c("#66C2A5", "#8DA0CB", "#FC8D62"),
    stroke_size = 0.5,
    show_percentage = F,
    set_name_size = 5
  )

all.venn

# ggsave(filename = "../Wytham_rodent_manuscript/revisions/new_figures/venn_virus_contigs_by_species_and_virus_groups_2022-06.pdf", contigs.venn)

```

# Virus genera

```{r}

get_venn_data_genera <-
  function(blast_results_table, contingency_table, list) {
    
    blast_contigs_subset <- blast_results_table %>%
      filter(`length` >= 50)
    
    contigency_table.s <-
      annotate_and_filter_contingency_table(contingency_table, blast_contigs_subset) %>%
      filter(virus_family %in% list) %>%
      rename(AS1 = WTCHG_714232_72815161_AS_1) %>%
      rename(AS2 = WTCHG_714232_72835163_AS_2) %>%
      rename(AS3 = WTCHG_714232_72865166_AS_3) %>%
      rename(AS4 = WTCHG_714232_71935169_AS_4) %>%
      rename(AS5 = WTCHG_714232_71965172_AS_5) %>%
      rename(MG1 = WTCHG_714232_72825162_MG_1) %>%
      rename(MG2 = WTCHG_714232_72845164_MG_2) %>%
      rename(MG3 = WTCHG_714232_72875167_MG_3) %>%
      rename(MG4 = WTCHG_714232_71945170_MG_4) %>%
      rename(MG5 = WTCHG_714232_71975173_MG_5) %>%
      rename(AF3 = WTCHG_714232_72855165_AF_3) %>%
      rename(AF4 = WTCHG_714232_72885168_AF_4) %>%
      rename(AF5 = WTCHG_714232_71955171_AF_5) %>%
      
      mutate(AS1_abundance =
               round(AS1 / ((subset(
                 samples_per_pool, Species == "AS" &
                   Interval == 1
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AS2_abundance =
               round(AS2 / ((subset(
                 samples_per_pool, Species == "AS" &
                   Interval == 2
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AS3_abundance =
               round(AS3 / ((subset(
                 samples_per_pool, Species == "AS" &
                   Interval == 3
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AS4_abundance =
               round(AS4 / ((subset(
                 samples_per_pool, Species == "AS" &
                   Interval == 4
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AS5_abundance =
               round(AS5 / ((subset(
                 samples_per_pool, Species == "AS" &
                   Interval == 5
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(MG1_abundance =
               round(MG1 / ((subset(
                 samples_per_pool, Species == "MG" &
                   Interval == 1
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(MG2_abundance =
               round(MG2 / ((subset(
                 samples_per_pool, Species == "MG" &
                   Interval == 2
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(MG3_abundance =
               round(MG3 / ((subset(
                 samples_per_pool, Species == "MG" &
                   Interval == 3
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(MG4_abundance =
               round(MG4 / ((subset(
                 samples_per_pool, Species == "MG" &
                   Interval == 4
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(MG5_abundance =
               round(MG5 / ((subset(
                 samples_per_pool, Species == "MG" &
                   Interval == 5
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AF3_abundance =
               round(AF3 / ((subset(
                 samples_per_pool, Species == "AF" &
                   Interval == 3
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AF4_abundance =
               round(AF4 / ((subset(
                 samples_per_pool, Species == "AF" &
                   Interval == 4
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      mutate(AF5_abundance =
               round(AF5 / ((subset(
                 samples_per_pool, Species == "AF" &
                   Interval == 5
               ))$`Number of paired-end reads`*2), digits=8)) %>%
      
      mutate(AS1 =
               ifelse((AS1_abundance <= (1 / 10000000)), 0, AS1)) %>%
      mutate(AS2 =
               ifelse((AS2_abundance <= (1 / 10000000)), 0, AS2)) %>%
      mutate(AS3 =
               ifelse((AS3_abundance <= (1 / 10000000)), 0, AS3)) %>%
      mutate(AS4 =
               ifelse((AS4_abundance <= (1 / 10000000)), 0, AS4)) %>%
      mutate(AS5 =
               ifelse((AS5_abundance <= (1 / 10000000)), 0, AS5)) %>%
      mutate(MG1 =
               ifelse((MG1_abundance <= (1 / 10000000)), 0, MG1)) %>%
      mutate(MG2 =
               ifelse((MG2_abundance <= (1 / 10000000)), 0, MG2)) %>%
      mutate(MG3 =
               ifelse((MG3_abundance <= (1 / 10000000)), 0, MG3)) %>%
      mutate(MG4 =
               ifelse((MG4_abundance <= (1 / 10000000)), 0, MG4)) %>%
      mutate(MG5 =
               ifelse((MG5_abundance <= (1 / 10000000)), 0, MG5)) %>%
      mutate(AF3 =
               ifelse((AF3_abundance <= (1 / 10000000)), 0, AF3)) %>%
      mutate(AF4 =
               ifelse((AF4_abundance <= (1 / 10000000)), 0, AF4)) %>%
      mutate(AF5 =
               ifelse((AF5_abundance <= (1 / 10000000)), 0, AF5)) %>%
      
      
      rowwise(`contig ID`) %>%
      mutate(total_AS = sum(
        c(
          AS1,
          AS2,
          AS3,
          AS4,
          AS5
        )
      )) %>%
      mutate(total_MG = sum(
        c(
          MG1,
          MG2,
          MG3,
          MG4,
          MG5
        )
      )) %>%
      mutate(total_AF = sum(
        c(
          AF3,
          AF4,
          AF5
        )
      )) %>%
      filter(sum(total_AS, total_AF, total_MG) >= 20) %>%
      filter(type != "NA")
    
    AS <- list(
      AS1 = rep(contigency_table.s$species_list, (contigency_table.s$AS1)),
      AS2 = rep(contigency_table.s$species_list, (contigency_table.s$AS2)),
      AS3 = rep(contigency_table.s$species_list, (contigency_table.s$AS3)),
      AS4 = rep(contigency_table.s$species_list, (contigency_table.s$AS4)),
      AS5 = rep(contigency_table.s$species_list, (contigency_table.s$AS5))
    )

    MG <- list(
      MG1 = rep(contigency_table.s$species_list, (contigency_table.s$MG1)),
      MG2 = rep(contigency_table.s$species_list, (contigency_table.s$MG2)),
      MG3 = rep(contigency_table.s$species_list, (contigency_table.s$MG3)),
      MG4 = rep(contigency_table.s$species_list, (contigency_table.s$MG4)),
      MG5 = rep(contigency_table.s$species_list, (contigency_table.s$MG5))
    )

    AF <- list(
      AF3 = rep(contigency_table.s$species_list, (contigency_table.s$AF3)),
      AF4 = rep(contigency_table.s$species_list, (contigency_table.s$AF4)),
      AF5 = rep(contigency_table.s$species_list, (contigency_table.s$AF5))
    )

    return(list(AS = AS, AF = AF, MG =  MG))
  }



get_virusgenera_venn <- function() {
  
  vert_viruses <-
    get_venn_data_genera(blast_results_table,
                         contingency_table,
                         unique(verteb_viruses.df$virus_family))
  invert_viruses <-
    get_venn_data_genera(blast_results_table,
                         contingency_table,
                         unique(invert_virus.df$virus_family))
  plant_viruses <-
    get_venn_data_genera(blast_results_table,
                         contingency_table,
                         unique(plant_virus.df$virus_family))
  bacterio_viruses <-
    get_venn_data_genera(blast_results_table,
                         contingency_table,
                         unique(bacteriophage.df$virus_family))
  
  vert.venn <-
    list(
      AS = unlist(vert_viruses$AS),
      AF = unlist(vert_viruses$AF),
      MG = unlist(vert_viruses$MG)
    ) %>%
    ggvenn(
      fill_color =  c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  vert.venn
  
  invert.venn <-
    list(
      AS = unlist(invert_viruses[[1]]),
      AF = unlist(invert_viruses[[2]]),
      MG = unlist(invert_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color =  c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  plant.venn <-
    list(
      AS = unlist(plant_viruses[[1]]),
      AF = unlist(plant_viruses[[2]]),
      MG = unlist(plant_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color =  c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  bacterio.venn <-
    list(
      AS = unlist(bacterio_viruses[[1]]),
      AF = unlist(bacterio_viruses[[2]]),
      MG = unlist(bacterio_viruses[[3]])
    ) %>%
    ggvenn(
      fill_color =  c("#66C2A5", "#8DA0CB", "#FC8D62"),
      stroke_size = 0.5,
      show_percentage = F,
      set_name_size = 5
    )
  
  virusgenera_venn <-
    (vert.venn | invert.venn) / (plant.venn | bacterio.venn)
  
  return(virusgenera_venn)
}

virusgenera_venn <- get_virusgenera_venn()

virusgenera_venn

# ggsave(filename = "../Wytham_rodent_manuscript/revisions/new_figures/venn_virus_genera_2022-06.pdf", virusgenera_venn)

```

# Species accumulation curve

```{r, echo=FALSE, warning=FALSE}

##### all viruses #####
all_virus_relabund <- get_relAbundance_s(unique(all_virus_species.df$virus_species), contingency_table.s)

# (unique(all_virus_relabund$virus_species))

all_virus_abund.df <- get_abundance(all_virus_relabund)

v_comm.mat <- get_comm_mat_s(all_virus_abund.df)

spec_accum <- specaccum(v_comm.mat[1:13,2:100],method = "random", permutations=100)

sp1 <- data.frame(species_richness = spec_accum$richness, sd = spec_accum$sd, samples = spec_accum$sites)

spec_accum_curve <-
  sp1 %>%
  ggplot(aes(samples, species_richness)) + 
  geom_line() + geom_point() +
  geom_ribbon(aes(ymin=species_richness-sd, ymax=species_richness+sd), alpha=0.2) + 
  theme_bw() + xlab("Sample number") + ylab("Virus richness") + ggtitle("All") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,100))

spec_accum_curve


##### vert viruses #####

vert_virus_relabund <- get_relAbundance_s(unique(verterb_virus_species.df$virus_species), contingency_table.s)

vert_virus_abund.df <- get_abundance(vert_virus_relabund)

vert_comm.mat <- get_comm_mat_s(vert_virus_abund.df)

spec_accum.vert <- specaccum(vert_comm.mat[1:13,2:21],method = "random", permutations=100)

sp.vert <- data.frame(species_richness = spec_accum.vert$richness, sd = spec_accum.vert$sd, samples = spec_accum.vert$sites)

spec_accum_curve.vert <-
  sp.vert  %>%
  ggplot(aes(samples, species_richness)) + 
  geom_line() + geom_point() +
  geom_ribbon(aes(ymin=species_richness-sd, ymax=species_richness+sd), alpha=0.2) + 
  theme_bw() + xlab("Sample number") + ylab("Virus richness") + ggtitle("Vertebrate-associated") + theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0,20))

spec_accum_curve.vert

##### bacteriophage #####

phage_relabund <- get_relAbundance_s(unique(phage_species.df$virus_species), contingency_table.s)

phage_abund.df <- get_abundance(phage_relabund)

phage_comm.mat <- get_comm_mat_s(phage_abund.df)

spec_accum.phage <- specaccum(phage_comm.mat[1:13,2:35],method = "random", permutations=100)

sp.phage <- data.frame(species_richness = spec_accum.phage$richness, sd = spec_accum.phage$sd, samples = spec_accum.phage$sites)

spec_accum_curve.phage <-
  sp.phage %>%
  ggplot(aes(samples, species_richness)) + 
  geom_line() + geom_point() +
  geom_ribbon(aes(ymin=species_richness-sd, ymax=species_richness+sd), alpha=0.2) + 
  theme_bw() + xlab("Sample number") + ylab("Virus richness") + ggtitle("Bacteriophage") + theme(plot.title = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,32), breaks=seq(0,32, 8 ))

spec_accum_curve.phage

spec_accum_plots <- ggarrange(spec_accum_curve, spec_accum_curve.vert,  spec_accum_curve.phage, common.legend = TRUE, nrow = 1)

spec_accum_plots

ggsave("../Wytham_rodent_manuscript/revisions/new_figures/rarefraction_curves_2022-06-09.pdf", spec_accum_plots, height=2.5, width=8)

```
## Picornavirus abundance

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=3, fig.height=2}

options(scipen = 2)

picorna_abundance_over_time <-
  
  get_relAbundance_s(unique(picorna$species_list), contingency_table.s)  %>% 
  mutate(Interval=as.character(Interval)) %>% 
  left_join(samples_per_pool) %>%
  mutate(viral_read_count =
           ifelse((viral_read_count/(`Number of paired-end reads`*2) <= (1/10000000)), 0, viral_read_count)) %>%
  # mutate(viral_read_count = 
  #        ifelse((viral_read_count < 5), 0, viral_read_count)) %>%
  mutate(Interval=as.character(Interval)) %>% 
  filter(virus_species %in% c("Mosavirus", 
                              "Cardiovirus", 
                              "Kunsagivirus", 
                              "Hunnivirus", 
                              "Unclassified picornavirus 1", 
                              "Unclassified picornavirus 2")) %>%
  group_by(virus_species, 
           Interval, 
           Species, 
           collection_date, 
           `Number of individuals`, 
           `Number of paired-end reads`) %>% 
  summarise(viral_read_sum = sum(viral_read_count)) %>%
  ggplot(aes(dmy(collection_date), 
             group=Species, colour = Species, 
             y = (viral_read_sum/`Number of individuals`/(`Number of paired-end reads`*2)))) +
  geom_point() + 
  geom_line() + 
  facet_wrap(~virus_species, nrow=2) + 
  scale_y_log10() +
  # scale_y_continuous(limits=c(-9, -4)) +
  ylab("Normalised read abundance") + 
  xlab("Time") + 
  scale_color_manual(values=c("#8DA0CB", "#66C2A5", "#FC8D62")) + 
  theme_bw() + 
  theme(legend.position="none")

picorna_abundance_over_time

ggsave("../Wytham_rodent_manuscript/revisions/new_figures/picornavirus_abundance_2022-06-09.pdf", picorna_abundance_over_time, height=4, width=6.5)

```

## Additive partitioning diversity

```{r}
library(vegan)

# for all three species
i1 <- c(seq(1,13,1)) # individual
i2 <- c(seq(3,5,1), rep(seq(1,5,1),2)) # time points
i3 <- c(rep(1,3), rep(2,5), rep(3,5)) # species
i4 <- rep(1,13) # total

# for individual species
i5 <- c(seq(1,5,1)) # individual
i7 <- rep(1,5) # total
i8 <- c(seq(3,5,1)) # individual
i9 <- rep(1,3) # total

levels <- data.frame(i1 = i1, i2 = i2, i4 = i4)
levels.2 <- data.frame(i5 = i5, i7 = i7)
levels.3 <- data.frame(i8 = i8, i9 = i9)

levels

v_comm_all.mat <- get_comm_mat_n(all_virus_abund.df)

v_comm_vert.mat <- get_comm_mat_n(vert_virus_abund.df)

v_comm_phage.mat <- get_comm_mat_n(phage_abund.df)

comm_mat.all <- v_comm_all.mat[1:13,2:100]
comm_mat.vert <- v_comm_vert.mat[1:13,2:21]
comm_mat.ph <- v_comm_phage.mat[1:13,2:35]

comm_mat_AF <- v_comm_all.mat[1:3,2:100]
comm_mat_AS <- v_comm_all.mat[4:8,2:100]
comm_mat_MG <- v_comm_all.mat[9:13,2:100]

adipart(comm_mat.all, levels, index = "richness", nsimul=99)
levels.2
```


# PERMANOVA
```{r}
#all viruses
set.seed(20220620)
comm_mat.all_virus <- all_virus_abund.df %>%
# filter(!(virus_family %in% c("Picobirnaviridae", "Microviridae", "Leviviridae"))) %>%
  get_comm_mat_s()


v_meta <- data.frame(species = c(rep("AF", 3), rep("AS", 5), rep("MG", 5)), 
                       interval = (c(3,4,5, rep(seq(1,5), 2))), id = comm_mat.all_virus$id)
v_meta$species <- as.factor(v_meta$species)
v_meta$species <- factor(v_meta$species, levels = c("MG", "AS", "AF"))

v_otu <-
comm_mat.all_virus[,2:length(comm_mat.all_virus[1,])] %>% 
  select_if(negate(function(col) is.numeric(col) && sum(col) == 0)) %>%
    select_if(negate(function(col) is.numeric(col) && sum(col) > 0.0005)) %>%
  as.matrix()


permanova <- adonis((v_otu) ~ species*interval,
                    data = v_meta, permutations=999, 
                    method = "bray",
                    strata = v_meta$species)

permanova$aov.tab

coefficients(permanova)
coef <- coefficients(permanova)["interval",]
top.coef <- coef[rev(order(abs(coef)))[1:20]]

top.coef.df <- data.frame(genera = names(top.coef), coef = as.numeric(top.coef))


hpermanova.all <- 
ggplot(top.coef.df,aes(x= reorder(genera,coef),coef))+geom_bar(stat ="identity") + coord_flip() + theme_minimal_vgrid() + xlab("Virus genera") + ylab("Coefficient")

hpermanova.all


# vertebrate viruses
set.seed(20220620)

comm_mat.vert <- vert_virus_abund.df %>%
  get_comm_mat_s()

v_otu <- comm_mat.vert[,2:length(comm_mat.vert[1,])] %>% 
  select_if(negate(function(col) is.numeric(col) && sum(col) == 0)) %>%
  select_if(negate(function(col) is.numeric(col) && sum(col) > 0.0005)) %>%
  as.matrix()


permanova <- adonis((v_otu) ~ interval*species,
                    data = v_meta, permutations=999, 
                    method = "bray",
                    strata = v_meta$species)

permanova$aov.tab

coef <- coefficients(permanova)["interval",]
top.coef <- coef[rev(order(abs(coef)))[1:10]]

top.coef.df <- data.frame(genera = names(top.coef), coef = as.numeric(top.coef))

hpermanova.vert <- 
ggplot(top.coef.df,aes(x= reorder(genera,coef),coef))+geom_bar(stat ="identity") + coord_flip() + theme_minimal_vgrid() + xlab("Virus genera") + ylab("Coefficient")

hpermanova.vert


# phage viruses
set.seed(20220620)

comm_mat.phage <- phage_abund.df %>%
  get_comm_mat_s()

v_otu <- comm_mat.phage[,2:length(comm_mat.phage[1,])] %>% 
  select_if(negate(function(col) is.numeric(col) && sum(col) == 0)) %>%
      select_if(negate(function(col) is.numeric(col) && sum(col) > 0.0005)) %>%
  as.matrix()

permanova <- adonis((v_otu) ~ interval*species,
                    data = v_meta, permutations=999, 
                    method = "bray", 
                    strata = v_meta$species)

permanova$aov.tab
coef <- coefficients(permanova)["interval",]
top.coef <- coef[rev(order(abs(coef)))[1:10]]

top.coef.df <- data.frame(genera = names(top.coef), coef = as.numeric(top.coef))

hpermanova.phage <- 
ggplot(top.coef.df,aes(x= reorder(genera,coef),coef))+geom_bar(stat ="identity") + coord_flip() + theme_minimal_vgrid() + xlab("Virus genera") + ylab("Coefficient")

hpermanova.phage

hpermonova.summary <-
ggarrange(hpermanova.all, hpermanova.vert, hpermanova.phage, labels = c("A", "B", "C"), nrow=1, ncol=3, widths = c(1,1,1), common.legend = TRUE)
hpermonova.summary
ggsave("../Wytham_rodent_manuscript/revisions/new_figures/hierarchical_permanova_2022-06.pdf", width = 18, height = 6, hpermonova.summary)


```

## Does the number indivs per sample pool impact the number of detected viruses?

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=4}

comm_mat <-
  all_virus_species.df %>%
  mutate(viral_read_count = ifelse(viral_read_count/(`Number of paired-end reads`*2) <= 1/10000000, 0, viral_read_count)) %>%
  get_comm_mat_s()

selected_table <-
  samples_per_pool %>%
  rename(Collection.date = collection_date) %>%
  select(Species, Interval, Collection.date, `Number of individuals`, id)

diversity_vs_sample_number <-
  get_diversity_stats(comm_mat) %>%
  rename(Shannon.diversity = H) %>%
  rename(Species.richness = species_richness) %>%
  rename(Collection.date = Interval) %>%
  left_join(selected_table) %>%
  rename(N_individuals = `Number of individuals`)


diversity_vs_sample_number %>%
  ggplot(aes(N_individuals, Species.richness)) + 
  geom_point() + 
  xlab("Number of individuals per pool") + 
  ylab("Number of viruses detected per pool")


result <- cor.test(diversity_vs_sample_number$N_individuals, diversity_vs_sample_number$Species.richness)

result

```

## Picornavirus diversity

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=4}

comm_mat <-
  all_virus_species.df %>%
   filter(virus_species %in% c("Cardiovirus", "Hunnivirus", "Kunsagivirus", "Mosavirus", "Unclassified picornavirus 1", "Unclassified picornavirus 2")) %>%
  mutate(viral_read_count = ifelse(viral_read_count/(`Number of paired-end reads`*2) <= 1/10000000, 0, viral_read_count)) %>%
  get_comm_mat_s()

diversity_stats <- get_diversity_stats(comm_mat) %>%
  rename(Shannon.diversity = H) %>%
  rename(Species.richness = species_richness) %>%
  rename(Collection.date = Interval)


diversity_stats$Species <- factor(diversity_stats$Species, levels=c("AS", "AF", "MG"))

p1 <- ggplot(diversity_stats, aes(dmy(Collection.date), Species.richness, group=Species, colour=Species))+
  geom_point() +
  geom_line() +
  facet_wrap(~Species) +
  scale_color_manual(values=c("#66C2A5", "#8DA0CB", "#FC8D62")) +
  theme_bw() +
  ylab("Richness")+
  scale_y_continuous(limits=c(0,5), breaks=seq(0,5,1))+
  # scale_x_date(limits =c(dmy("01/01/2017"), dmy("31/12/2017")))+
  # theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none")+
  xlab("Time")

p2 <- ggplot(diversity_stats, aes(dmy(Collection.date), Shannon.diversity, group=Species, colour=Species))+
  geom_point() +
  geom_line() +
  facet_wrap(~Species) +
  scale_color_manual(values=c("#66C2A5", "#8DA0CB", "#FC8D62")) +
  theme_bw()+
  ylab("Eveness") +
  scale_y_continuous(limits=c(0,1.0), breaks=seq(0,1,0.2))+
  # scale_x_date(limits =c(dmy("01/01/2017"), dmy("31/12/2017")))+
  # theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none")+
  xlab("Time")

picorna.fig <- ggarrange(p2, p1, align="hv", ncol=1, legend = "none")
picorna.fig <- annotate_figure(picorna.fig)
picorna.fig
ggsave("../Wytham_rodent_manuscript/revisions/new_figures/picornavirus_diversity_2022-06.pdf", picorna.fig, height=4, width=5.5)
 picorna.fig

```



```{r}

####################
# Time series of viral diversity and
# environmental predictors

######################
# required libraries
library(mgcv)
library(lubridate)
library(dplyr)
library(tidyverse)
library(patchwork)


diversity_stats$date <- as.Date(diversity_stats$Collection.date, "%d/%m/%Y")
diversity_stats$dec_date <- decimal_date(diversity_stats$date)

AS_diversity <-
  diversity_stats %>%
  filter(Species %in% c("AS"))
head(AS_diversity)
MG_diversity <-
  diversity_stats %>%
  filter(Species %in% c("MG"))
head(MG_diversity)

date_range = range(diversity_stats$dec_date)
pred_dates_div <- seq(date_range[1],date_range[2]+(2*1/(365/14)),by = 1/(365/14))

# Impute Shannon diversity
impute_AS_div <- loess(Shannon.diversity~dec_date, data = AS_diversity, span = 0.75)
impute_MG_div <- loess(Shannon.diversity~dec_date, data = MG_diversity, span = 0.75)

impute_AS_df <- predict(impute_AS_div, data.frame(dec_date = pred_dates_div))
impute_MG_df <- predict(impute_MG_div, data.frame(dec_date = pred_dates_div))

# Impute Species richness

impute_AS_rich <- loess(Species.richness~dec_date, data = AS_diversity, span = 0.75)
impute_MG_rich <- loess(Species.richness~dec_date, data = MG_diversity, span = 0.75)

impute_AS_rich_df <- predict(impute_AS_rich, data.frame(dec_date = pred_dates_div))
impute_MG_rich_df <- predict(impute_MG_rich, data.frame(dec_date = pred_dates_div))

# Imputed  vs observed
plot(pred_dates_div, impute_AS_df, col = '#33a8b5', 
     ylim = c(0,0.9), type = 'b', xlim = c(2017,2018),
     xlab = 'decimal dates', ylab = 'shannon diversity')
points(AS_diversity$dec_date, AS_diversity$Shannon.diversity, col = '#33a8b5', pch = 19)
points(pred_dates_div, impute_MG_df, col = '#dea310', type = 'b')
points(MG_diversity$dec_date, MG_diversity$Shannon.diversity, col = '#dea310', pch = 19)
legend(2017.8,0.8, legend=c("AS", "MG", 'observed', 'imputed'),
       col=c("#33a8b5", "#dea310",'black','black'), pch=c(19,19,19,1), cex=0.8, bty = 'n')

```

```{r}

######### Combining Sawmill and Upper seeds #########

sawmill1 = read.csv('../Wytham_rodent_manuscript/cross-correlation_analysis/wytham_sawmill_2017.csv')
upperseed = read.csv('../Wytham_rodent_manuscript/cross-correlation_analysis/wytham_upperseeds_2017.csv')

sawmill1$station <- rep("sawmill", length(sawmill1$date_format))
upperseed$station <- rep("upperseed", length(upperseed$date_format))

# hum sensors at sawmill reading too high, so excluding
sawmill1$hum. <- rep(NA, length(sawmill1$hum.))

weather_combined <- rbind(sawmill1, upperseed)

weather_combined$date_format =as.POSIXct(weather_combined$date_format, 
                                         tz = "GMT", format = '%d/%m/%Y %H:%M')

weather_combined$date2 =format(weather_combined$date_format, format = '%d/%m/%Y')
weather_combined = weather_combined[!(is.na(weather_combined$date2)),]

weather_combined_trun = weather_combined[,c('date2', 'rain_cts','hum.','t_hu')]


```

```{r}
## daily 
weather_daily_combined  <- weather_combined_trun %>%
  group_by(date2) %>% 
  dplyr::summarise(rain_cts_cum = sum(rain_cts),
                   rain_cts_mean = mean(rain_cts, na.rm = TRUE),
                   hum_mean = mean(hum., na.rm = TRUE),
                   hum_sd = sd(hum., na.rm = TRUE),
                   hum_min = min(hum., na.rm = TRUE),
                   hum_max = max(hum., na.rm = TRUE),
                   temp_mean = mean(t_hu, na.rm = TRUE),
                   temp_sd = sd(t_hu, na.rm = TRUE),
                   temp_min = min(t_hu, na.rm = TRUE),
                   temp_max = max(t_hu, na.rm = TRUE))

weather_daily_combined$dec_date <- decimal_date(as.Date(weather_daily_combined$date2, format = '%d/%m/%Y'))

dates_weather = seq((date_range[1]-(15*1/(365/14))),date_range[2]+(2*1/(365/14)),by = 1/(365/14))

# temperature

temp_mean_gam <- gam(formula = temp_mean~s(dec_date, bs="cc"), data = weather_daily_combined)
temp_mean_pred <- predict(temp_mean_gam, data.frame(dec_date = dates_weather))
temp_min_gam <- gam(formula = temp_min~s(dec_date, bs="cc"), data = weather_daily_combined)
temp_min_pred <- predict(temp_min_gam, data.frame(dec_date = dates_weather))
temp_max_gam <- gam(formula = temp_max~s(dec_date, bs="cc"), data = weather_daily_combined)
temp_max_pred <- predict(temp_max_gam, data.frame(dec_date = dates_weather))

plot(dates_weather, temp_mean_pred, pch = 19, ylim = c(0,25))
points(dates_weather, temp_max_pred, col = 'darkred')
points(dates_weather, temp_min_pred, col = 'navy')

#humidity
hum_gam <- gam(formula = hum_mean~s(dec_date, bs="cc"), data = weather_daily_combined)
hum_pred <- predict(hum_gam, data.frame(dec_date = dates_weather))
plot(dates_weather, hum_pred, pch = 19, type = 'b', 
     main = 'humidity predictions', ylim = c(70,100))
points(weather_daily_combined$dec_date, weather_daily_combined$hum_mean, col = 'grey80')

# Rain
rain_gam <- gam(formula = rain_cts_cum~s(dec_date, bs="cc"), data = weather_daily_combined)
rain_pred <- predict(rain_gam, data.frame(dec_date = dates_weather))
plot(dates_weather, rain_pred, pch = 19, type = 'b', 
     main = 'rainfall predictions', ylim = c(3,8))
points(weather_daily_combined$dec_date, weather_daily_combined$rain_cts_cum, col = 'grey80')

```


```{r}
#### weekly summaries...
year_week <- function(x,base) week(x) - week(base) + 52*(year(x) - year(base))

class(weather_daily_combined$date2 )
weather_daily_combined$date2 = as.Date(weather_daily_combined$date2, format = '%d/%m/%Y')
weather_daily_combined$week = year_week(weather_daily_combined$date2, "2016-01-01")
weather_daily_combined$week[order(weather_daily_combined$week)]
names(weather_daily_combined)

weather_weekly_combined  <- weather_daily_combined %>% 
  group_by(week) %>% 
  dplyr::summarise(min_date = min(date2),
                   rain_cts_cum_week = sum(rain_cts_cum, na.rm = TRUE),
                   rain_cts_mean_daily = mean(rain_cts_cum, na.rm = TRUE),
                   hum_cum_week = sum(hum_mean, na.rm = TRUE),
                   hum_mean_daily =  mean(hum_mean, na.rm = TRUE),
                   hum_max_daily =  mean(hum_max, na.rm = TRUE),
                   hum_min_daily =  mean(hum_min, na.rm = TRUE),
                   
                   temp_cum_week = sum(temp_mean, na.rm = TRUE),
                   temp_mean_daily =  mean(temp_mean, na.rm = TRUE),
                   temp_max_daily =  mean(temp_max, na.rm = TRUE),
                   temp_min_daily =  mean(temp_min, na.rm = TRUE))

weather_weekly_combined$dec_date  = decimal_date(weather_weekly_combined$min_date)
weather_weekly_combined$dec_date = weather_weekly_combined$dec_date +1/(365/3)

rain_wk_gam <- gam(formula = rain_cts_cum_week~s(dec_date, bs="cc"), data = weather_weekly_combined)
rain_wk_pred <- predict(rain_wk_gam, data.frame(dec_date = dates_weather))
plot(dates_weather, rain_wk_pred, pch = 19, type = 'b', 
     main = 'rainfall predictions', ylim = c(0,80))
points(weather_weekly_combined$dec_date, weather_weekly_combined$rain_cts_cum_week, col = 'grey80')



# Host pop density

MNKA = read.csv('../Wytham_rodent_manuscript/cross-correlation_analysis/MNKA_wytham_since_Nov2016.csv', header = TRUE)
pop_density <- 
  MNKA %>%
  mutate(collection_date = (Collection_date)) %>%
  select(Species, collection_date, MNKA_per_ha)

pop_density$collection_date = as.Date(pop_density$collection_date, format = '%d/%m/%Y')
pop_density$dec_date = decimal_date(pop_density$collection_date)

AS_density <-
  pop_density %>%
  filter(Species %in% c("AS"))
MG_density <-
  pop_density %>%
  filter(Species %in% c("MG"))

ASden_gam <- loess(formula = MNKA_per_ha~(dec_date), data = AS_density, span = 0.3)
ASden_pred <- predict(ASden_gam, data.frame(dec_date = dates_weather))

MGden_gam <- loess(formula = MNKA_per_ha~(dec_date), data = MG_density, span = 0.3)
MGden_pred <- predict(MGden_gam, data.frame(dec_date = dates_weather))

```

# Comparing imputed and observed data 

```{r}
############
# combining dataframes
diversity <- data.frame(dec_dates = round(pred_dates_div,3),
                        AS_diversity = impute_AS_df,
                        MG_diversity = impute_MG_df,
                        AS_rich = impute_AS_rich_df,
                        MG_rich = impute_MG_rich_df)

weather <- data.frame(dec_dates = round(dates_weather, 3), # have to round to match
                      temp_mean = temp_mean_pred,
                      temp_min = temp_min_pred,
                      temp_max = temp_max_pred,
                      hum = hum_pred,
                      rain = rain_pred,
                      rain_wk = rain_wk_pred,
                      AS_density = ASden_pred,
                      MG_density = MGden_pred)

timeseries_df = merge(diversity,weather, by ='dec_dates', all = TRUE )

write_csv(timeseries_df, "../Wytham_rodent_manuscript/revisions/new_figures/imputed_timeseries_data_Wytham_rodent_virome_2022-06.csv")

## plotting imputed and observed values for diversity and density 

imputed_data <-
  timeseries_df %>%
  select(dec_dates, AS_diversity, MG_diversity, AS_rich, MG_rich, AS_density, MG_density)%>%
  mutate(dec_dates = as.Date(date_decimal(dec_dates))) %>%
  rename(Collection.date = dec_dates) %>%
  melt(id.var=c("Collection.date")) %>%
  separate(variable, sep = "_", into = c("Species", "variable")) %>%
  mutate(variable = ifelse(variable == "diversity", "Viral evenness", variable)) %>%
  mutate(variable = ifelse(variable == "rich", "Viral richness", variable)) %>%
  mutate(variable = ifelse(variable == "density", "MNKA_per_ha", variable)) %>%
  select(Collection.date, Species, variable, value) %>%
  mutate(Type = rep("Imputed", length(Collection.date)))


observed_data <-
  diversity_stats %>%
  rename(`Viral evenness` = Shannon.diversity) %>%
  rename(`Viral richness` = Species.richness) %>%
  select(Collection.date, `Viral evenness`, `Viral richness`, Species) %>%
  filter(Species %in% c("AS", "MG")) %>%
  melt(id.var=c("Collection.date", "Species")) %>%
  combine(
    MNKA %>%
      select(Collection_date, MNKA_per_ha, Species) %>%
      filter(Species %in% c("AS", "MG")) %>%
      melt(id.vars=c("Collection_date", "Species")) %>%
      rename(Collection.date = Collection_date)
  ) %>%
  mutate(Type = rep("Observed", length(Collection.date))) %>%
  mutate(Collection.date = dmy((Collection.date)))


combined_df <- rbind(imputed_data, observed_data)

combined_df$variable <- factor(combined_df$variable, levels=c("Viral evenness", "Viral richness", "MNKA_per_ha"))

fitted_vs_observed <- 
  combined_df %>%
  ggplot(aes(Collection.date, value, group = variable, alpha = Type, shape = Type)) +
  geom_point(size=2) +
  facet_wrap(Species~variable, scales = "free") + 
  scale_shape_manual(values = c(21, 16)) + 
  scale_alpha_manual(values=c(1,0.7)) + 
  theme_bw() + 
  ylab("Value") + xlab("Time") + theme(legend.position = "top")


weather_plot <- 
  timeseries_df %>%
  select(dec_dates, temp_mean, hum, rain)%>%
  mutate(dec_dates = as.Date(date_decimal(dec_dates))) %>%
  rename(Collection.date = dec_dates) %>%
  melt(id.var=c("Collection.date")) %>%
  select(Collection.date, variable, value) %>%
  ggplot(aes(Collection.date, value,group = variable)) + 
  geom_point(colour = "blue") + 
  facet_wrap(~variable, scales="free_y") + 
  theme_bw()  + 
  ylab("Value") + xlab("Time")

weather_plot

combined_plot <- ggarrange(fitted_vs_observed, weather_plot,heights = c(1, 0.6), nrow = 2, labels = c("A", "B"))
combined_plot
ggsave("../Wytham_rodent_manuscript/revisions/new_figures/Figure_SX_fitted_vs_observed_host_species_metrics.pdf", combined_plot, height = 8, width=10)


```

# Cross correlation analysis 

```{r}

########################
# cross correlaton function

names(timeseries_df)
par(mfrow=c(3,2))
ccf_values_AS1 <- ccf(timeseries_df$AS_diversity,timeseries_df$temp_mean, 
                      na.action = na.pass, lag.max = 12, main = 'mean temp', col ='#33a8b5')
# ccf(timeseries_df$AS_diversity,timeseries_df$temp_min, 
#     na.action = na.pass, lag.max = 6)
ccf_values_AS2 <- ccf(timeseries_df$AS_diversity,timeseries_df$temp_max, 
                      na.action = na.pass, lag.max = 12,  main ='max temp', col ='#33a8b5')
ccf_values_AS3 <- ccf(timeseries_df$AS_diversity,timeseries_df$hum, 
                      na.action = na.pass, lag.max = 12,  main = 'humidity', col ='#33a8b5')
ccf_values_AS4 <- ccf(timeseries_df$AS_diversity,timeseries_df$rain, 
                      na.action = na.pass, lag.max = 12, main = 'rainfall', col ='#33a8b5')
ccf_values_AS5 <- ccf(timeseries_df$AS_diversity,timeseries_df$rain_wk, 
                      na.action = na.pass, lag.max = 12,main = 'weekly rainfall', col ='#33a8b5')
ccf_values_AS6 <- ccf(timeseries_df$AS_diversity,timeseries_df$AS_density, 
                      na.action = na.pass, lag.max = 12, main = 'density', col ='#33a8b5')

ccf_values_AS1 # temp mean; max sig lag = -7 (0.640)
ccf_values_AS2 # temp max; max sig lag = -7 (0.629)
ccf_values_AS3 # hum; max sig lag = -5 (0.630)
ccf_values_AS4 # rain; max sig lag = 0 (-0.385)
ccf_values_AS5 # rain_wk; max sig lag = 0 (-0.699)
ccf_values_AS6 # density; max sig lag = -6 (-0.418)

par(mfrow=c(3,2))
ccf_values_MG1 <- ccf(timeseries_df$MG_diversity,timeseries_df$temp_mean, 
                      na.action = na.pass, lag.max = 12,  main = 'mean temp', col ='#dea310')
# ccf(timeseries_df$MG_diversity,timeseries_df$temp_min, 
#     na.action = na.pass, lag.max = 6)
ccf_values_MG2 <- ccf(timeseries_df$MG_diversity,timeseries_df$temp_max, 
                      na.action = na.pass, lag.max = 12,  main = 'max temp', col = '#dea310')
ccf_values_MG3 <- ccf(timeseries_df$MG_diversity,timeseries_df$hum, 
                      na.action = na.pass, lag.max = 12,  main = 'humidity', col = '#dea310')
ccf_values_MG4 <- ccf(timeseries_df$MG_diversity,timeseries_df$rain, 
                      na.action = na.pass, lag.max = 12,main = ' rainfall', col = '#dea310')
ccf_values_MG5 <- ccf(timeseries_df$MG_diversity,timeseries_df$rain_wk, 
                      na.action = na.pass, lag.max = 12,main = 'weekly rainfall', col = '#dea310')
ccf_values_MG6 <- ccf(timeseries_df$MG_diversity,timeseries_df$MG_density, 
                      na.action = na.pass, lag.max = 12,main = 'density', col = '#dea310')

ccf_values_MG1 # temp mean; max sig lag = -3 (0.743)
ccf_values_MG2 # temp max; max sig lag = -2 (0.749)
ccf_values_MG3 # hum; max sig lag = 0 (-0.767)
ccf_values_MG4 # rain; max sig lag = -7 (-0.407)
ccf_values_MG5 # rain_wk; max sig lag = -7 (-0.465)
ccf_values_MG6 # density; max sig lag = -5 (-0.715)

par(mfrow=c(3,2))
ccf_values_rich_AS1 <- ccf(timeseries_df$AS_rich,timeseries_df$temp_mean, 
                           na.action = na.pass, lag.max = 12, main = 'mean temp', col ='#33a8b5')
# ccf(timeseries_df$AS_diversity,timeseries_df$temp_min, 
#     na.action = na.pass, lag.max = 6)
ccf_values_rich_AS2 <- ccf(timeseries_df$AS_rich,timeseries_df$temp_max, 
                           na.action = na.pass, lag.max = 12,  main ='max temp', col ='#33a8b5')
ccf_values_rich_AS3 <- ccf(timeseries_df$AS_rich,timeseries_df$hum, 
                           na.action = na.pass, lag.max = 12,  main = 'humidity', col ='#33a8b5')
ccf_values_rich_AS4 <- ccf(timeseries_df$AS_rich,timeseries_df$rain, 
                           na.action = na.pass, lag.max = 12, main = 'rainfall', col ='#33a8b5')
ccf_values_rich_AS5 <- ccf(timeseries_df$AS_rich,timeseries_df$rain_wk, 
                           na.action = na.pass, lag.max = 12,main = 'weekly rainfall', col ='#33a8b5')
ccf_values_rich_AS6 <- ccf(timeseries_df$AS_rich,timeseries_df$AS_density, 
                           na.action = na.pass, lag.max = 12, main = 'density', col ='#33a8b5')

ccf_values_rich_AS1 # temp_mean; max sig lag = -1 (0.829) 0 (0.770)
ccf_values_rich_AS2 # temp_max; max sig lag = -1 (0.826) 0 (0.790)
ccf_values_rich_AS3 # hum; max sig lag = 0 (-0.891) 0 (-0.759)
ccf_values_rich_AS4 # rain; max sig lag = -5 (0.401) -4 (0.393)
ccf_values_rich_AS5 # rain_wk; max sig lag = -6 (0.527) -6 (0.506)
ccf_values_rich_AS6 # density; max sig lag = 0 (-0.519) 0 (-0.203)

par(mfrow=c(3,2))
ccf_values_rich_MG1 <- ccf(timeseries_df$MG_rich,timeseries_df$temp_mean, 
                           na.action = na.pass, lag.max = 12, main = 'mean temp', col ='#dea310')
# ccf(timeseries_df$AS_diversity,timeseries_df$temp_min, 
#     na.action = na.pass, lag.max = 6)
ccf_values_rich_MG2 <- ccf(timeseries_df$MG_rich,timeseries_df$temp_max, 
                           na.action = na.pass, lag.max = 12,  main ='max temp', col ='#dea310')
ccf_values_rich_MG3 <- ccf(timeseries_df$MG_rich,timeseries_df$hum, 
                           na.action = na.pass, lag.max = 12,  main = 'humidity', col ='#dea310')
ccf_values_rich_MG4 <- ccf(timeseries_df$MG_rich,timeseries_df$rain, 
                           na.action = na.pass, lag.max = 12, main = 'rainfall', col ='#dea310')
ccf_values_rich_MG5 <- ccf(timeseries_df$MG_rich,timeseries_df$rain_wk, 
                           na.action = na.pass, lag.max = 12,main = 'weekly rainfall', col ='#dea310')
ccf_values_rich_MG6 <- ccf(timeseries_df$MG_rich,timeseries_df$AS_density, 
                           na.action = na.pass, lag.max = 12, main = 'density', col ='#dea310')

ccf_values_rich_MG1 # temp_mean; max sig lag = -2 (0.773) -1 0.801
ccf_values_rich_MG2 # temp_max; max sig lag = -2 (0.775) -1 (0.792)
ccf_values_rich_MG3 # hum; max sig lag = 0 (-0.820) 0 (-0.883)
ccf_values_rich_MG4 # rain; max sig lag = -6 (0.403) -5 (0.381)
ccf_values_rich_MG5 # rain_wk; max sig lag = -7 (0.494) -7 (0.503)
ccf_values_rich_MG6 # density; max sig lag = 0 (-0.409) 0 (-0.605)

##########
#wide to long
library(reshape2)
library(ggplot2)
names(timeseries_df)
timeseries_long = melt(timeseries_df, id.vars=c("dec_dates"))#,'AS_diversity', 'MG_diversity','AS_rich', 'MG_rich'))
head(timeseries_long)
ggplot(timeseries_long, aes(dec_dates,value))+
  geom_line()+
  #geom_point(aes(dec_dates,AS_diversity), col = '#33a8b5')+
  facet_wrap(.~variable, scales = 'free_y')+
  theme_classic()

```

```{r}

###### Check correlations among predictors

library(ggcorrplot)

ccf_values_rich_AS1 # temp_mean; max sig lag =  0 (0.770)
ccf_values_rich_AS2 # temp_max; max sig lag = 0 (0.790)
ccf_values_rich_AS3 # hum; max sig lag =  0 (-0.759)
ccf_values_rich_AS4 # rain; max sig lag = -4 (0.393)
ccf_values_rich_AS5 # rain_wk; max sig lag = -6 (0.506)
ccf_values_rich_AS6 # density; max sig lag = 0 (-0.203)

ccf_values_rich_MG1 # temp_mean; max sig lag = -1 0.801
ccf_values_rich_MG2 # temp_max; max sig lag = -1 (0.792)
ccf_values_rich_MG3 # hum; max sig lag = 0 (-0.883)
ccf_values_rich_MG4 # rain; max sig lag = -5 (0.381)
ccf_values_rich_MG5 # rain_wk; max sig lag = -7 (0.503)
ccf_values_rich_MG6 # density; max sig lag = 0 (-0.605)

ccf_values_AS1 # temp mean; max sig lag = -7 (0.639)
ccf_values_AS2 # temp max; max sig lag = -7 (0.629)
ccf_values_AS3 # hum; max sig lag = -5 (0.630)
ccf_values_AS4 # rain; max sig lag = 0 (-0.386)
ccf_values_AS5 # rain_wk; max sig lag = 0 (-0.699)
ccf_values_AS6 # density; max sig lag = -6 (-0.419)

ccf_values_MG1 # temp mean; max sig lag = -2 (0.750)
ccf_values_MG2 # temp max; max sig lag = -2 (0.751)
ccf_values_MG3 # hum; max sig lag = 0 (-0.786)
ccf_values_MG4 # rain; max sig lag = -7 (-0.404)
ccf_values_MG5 # rain_wk; max sig lag = -7 (-0.468)
ccf_values_MG6 # density; max sig lag = -4 (-0.720)

predictor_df <- data.frame(AS_diversity = timeseries_df$AS_diversity,
                           MG_diversity = timeseries_df$MG_diversity,
                           AS_rich = timeseries_df$AS_rich,
                           MG_rich = timeseries_df$MG_rich,
                           temp = lag(timeseries_df$temp_mean, 0),
                           temp_1 = lag(timeseries_df$temp_mean, 1),
                           temp_2 = lag(timeseries_df$temp_mean, 2),
                           temp_3 = lag(timeseries_df$temp_mean, 3),
                           temp_7 = lag(timeseries_df$temp_mean, 7),
                           hum = lag(timeseries_df$hum, 0),
                           hum_5 = lag(timeseries_df$hum, 5),
                           rain = lag(timeseries_df$rain, 0),    
                           rain_4 = lag(timeseries_df$rain, 4),
                           rain_5 = lag(timeseries_df$rain, 5),
                           # rain_6 = lag(timeseries_df$rain, 6),
                           rain_7 = lag(timeseries_df$rain, 7),
                           AS_density = lag(timeseries_df$AS_diversity, 0),
                           AS_density_6 = lag(timeseries_df$AS_diversity, 6), 
                           MG_density = lag(timeseries_df$MG_diversity, 0),
                          MG_density_4 = lag(timeseries_df$MG_diversity, 4),
                           MG_density_5 = lag(timeseries_df$MG_diversity, 5))


# Create four set of predictors corresponding to each tested model

AS_diversity_predictors <- data.frame(temp_7 = predictor_df$temp_7, 
                                      hum_5 = predictor_df$hum_5,
                                      rain = predictor_df$rain,
                                      AS_density_6 = predictor_df$AS_density_6)


MG_diversity_predictors <- data.frame(temp_2 = predictor_df$temp_2, 
                                      hum = predictor_df$hum,
                                      rain_7 = predictor_df$rain_7,
                                      MG_density_4 = predictor_df$MG_density_4)


AS_rich_predictors <- data.frame(temp = predictor_df$temp, 
                                 hum = predictor_df$hum,
                                 rain_4 = predictor_df$rain_4,
                                 AS_density = predictor_df$AS_density)


MG_rich_predictors <- data.frame(temp_1 = predictor_df$temp_1, 
                                 hum = predictor_df$hum,
                                 rain_5 = predictor_df$rain_5,
                                 MG_density = predictor_df$MG_density)

AS_diversity_M = cor(na.omit(AS_diversity_predictors))
MG_diversity_M = cor(na.omit(MG_diversity_predictors))
AS_rich_M = cor(na.omit(AS_rich_predictors))
MG_rich_M = cor(na.omit(MG_rich_predictors))

corr_AS_d <- ggcorrplot(AS_diversity_M, method = 'square', type = "lower", lab = TRUE) + theme_nothing() #(rain+hum)
corr_MG_d <- ggcorrplot(MG_diversity_M, method = 'square', type = "lower", lab = TRUE) + theme_nothing() #(rain+hum) + (density+temp)
corr_AS_r <- ggcorrplot(AS_rich_M, method = 'square', type = "lower", lab = TRUE) + theme_nothing() # temp+hum + rain+density
corr_MG_r <-  ggcorrplot(MG_rich_M, method = 'square', type = "lower", lab = TRUE) + theme_nothing() # temp+hum + hum+density

corr_AS_r  + theme_minimal()

corrplot_summ <- ggarrange(corr_AS_d, corr_MG_d, corr_AS_r, corr_MG_r, labels=c("Eveness - AS", "Eveness - MG", "Richness - AS", "Richness - MG"), nrow = 2, ncol = 2, label.x = -0.1, label.y = 1)

ggsave("../Wytham_rodent_manuscript/revisions/new_figures/corrplot_summary_2022-06.pdf", corrplot_summ, width = 6, height = 7)
# AS diversity -> hum_5 and rain are correlated (-0.93)
# temp + hum + density
# temp + rain + density
# temp + hum
# temp + density
# temp + rain
# rain + density
# density + hum

# MG diversity -> temp_3 + density_5 correlated (0.89) and hum and density correlated (0.7)
# temp + hum
# temp + rain
# rain + density
# rain + hum
# rain + temp + hum


# AS rich -> rain and density correlated (-0.72) + temp+hum (-0.88)
# temp + rain
# temp + density
# hum + density
# rain + hum


# MG rich -> hum and density correlated (-0.84) + temp+hum (-0.9)
# temp + rain
# temp + density
# rain + hum
# rain + density
# temp + density + rain



```

```{r}

#########
# GLM


AS_div_model_a <- glm(data = predictor_df, formula = AS_diversity ~ temp_7 + hum_5 + AS_density_6, family = gaussian)
AS_div_model_b <- glm(data = predictor_df, formula = AS_diversity ~ temp_7 + rain + AS_density_6, family = gaussian)
AS_div_model_c <- glm(data = predictor_df, formula = AS_diversity ~ temp_7 + hum_5, family = gaussian)
AS_div_model_d <- glm(data = predictor_df, formula = AS_diversity ~ temp_7 + AS_density_6, family = gaussian) ###
AS_div_model_e <- glm(data = predictor_df, formula = AS_diversity ~ rain + temp_7, family = gaussian)
AS_div_model_f <- glm(data = predictor_df, formula = AS_diversity ~ rain + AS_density_6, family = gaussian)
AS_div_model_g <- glm(data = predictor_df, formula = AS_diversity ~ hum_5 + AS_density_6, family = gaussian)


#MG_diversity 

MG_div_model_a <- glm(data = predictor_df, formula = MG_diversity ~ temp_2 + hum, family = gaussian)
MG_div_model_b <- glm(data = predictor_df, formula = MG_diversity ~ temp_2 + rain, family = gaussian) ###
MG_div_model_c <- glm(data = predictor_df, formula = MG_diversity ~ rain_7 + MG_density_4, family = gaussian)
MG_div_model_d <- glm(data = predictor_df, formula = MG_diversity ~ rain_7 + hum, family = gaussian)
MG_div_model_e <- glm(data = predictor_df, formula = MG_diversity ~ rain_7 + hum + temp_2, family = gaussian)

# AS rich -> rain and density correlated (-0.72) + temp+hum (-0.88)
# temp + rain
# temp + density
# hum + density
# rain + hum


#AS_rich 

AS_rich_model_a <- glm(data = predictor_df, formula = AS_rich ~ temp + rain_4, family = gaussian)
AS_rich_model_b <- glm(data = predictor_df, formula = AS_rich ~ rain_4 + hum, family = gaussian) ###
AS_rich_model_c <- glm(data = predictor_df, formula = AS_rich ~ temp + AS_density, family = gaussian)
AS_rich_model_d <- glm(data = predictor_df, formula = AS_rich ~ hum + AS_density, family = gaussian)



# MG rich -> hum and density correlated (-0.84) + temp+hum (-0.9)
# temp + rain
# temp + density
# rain + hum
# rain + density
# temp + density + rain

#MG_rich 

MG_rich_model_a <- glm(data = predictor_df, formula = MG_rich ~ temp_1 + rain_5, family = gaussian)
MG_rich_model_b <- glm(data = predictor_df, formula = MG_rich ~ temp_1 + MG_density + rain_5, family = gaussian) ###
MG_rich_model_c <- glm(data = predictor_df, formula = MG_rich ~ temp_1 + MG_density, family = gaussian)
MG_rich_model_d <- glm(data = predictor_df, formula = MG_rich ~ rain_5 + hum, family = gaussian)
MG_rich_model_e <- glm(data = predictor_df, formula = MG_rich ~ rain_5 + MG_density, family = gaussian)




library(AICcmodavg)
# AS diversity (model d) - temp + density
AS_div_models <- list(AS_div_model_a, AS_div_model_b, 
                      AS_div_model_c, AS_div_model_d, 
                      AS_div_model_e, AS_div_model_f, AS_div_model_g)
model.names <- c("model_a", "model_b", "model_c", "model_d", "model_e", "model_f", "model_g")
aictab(cand.set = AS_div_models, modnames = model.names)

# AS richness (model c) - temp + density
AS_rich_models <- list(AS_rich_model_a, AS_rich_model_b, 
                       AS_rich_model_c, AS_rich_model_d)
model.names <- c("model_a", "model_b", "model_c", "model_d")
aictab(cand.set = AS_rich_models, modnames = model.names)

# MG diversity (model e) - rain + hum + temp
MG_div_models <- list(MG_div_model_a, MG_div_model_b, 
                      MG_div_model_c, MG_div_model_d, MG_div_model_e)
model.names <- c("model_a", "model_b", "model_c", "model_d", "model_e")
aictab(cand.set = MG_div_models, modnames = model.names)

# MG richness (model c) - rain + hum
MG_rich_models <- list(MG_rich_model_a, MG_rich_model_b, 
                       MG_rich_model_c, MG_rich_model_d, 
                       MG_rich_model_e)
model.names <- c("model_a", "model_b", "model_c", "model_d", "model_e")
aictab(cand.set = MG_rich_models, modnames = model.names)


####### Plotting GLM results ######

library(jtools)
library(broom.mixed)
library(reghelper)
library(ggpubr)

AS_div_fit <- AS_div_model_d
MG_div_fit <- MG_div_model_e
AS_rich_fit <- AS_rich_model_c
MG_rich_fit <- MG_rich_model_c

beta(AS_rich_fit)
export_summs(AS_div_fit, MG_div_fit, AS_rich_fit, MG_rich_fit, scale = TRUE, transform.response = TRUE, error_format = "[{conf.low}, {conf.high}]")
summary(AS_div_fit)

sd <- plot_summs(AS_div_fit, MG_div_fit, scale = TRUE) + 
  scale_color_manual(values=c("#FC8D62", "#66C2A5")) + 
  scale_shape_manual(values=c("circle", "circle")) + theme_minimal()
sr <- plot_summs(AS_rich_fit, MG_rich_fit, scale = TRUE)  + 
  scale_color_manual(values=c("#FC8D62", "#66C2A5")) + 
  scale_shape_manual(values=c("circle", "circle")) + theme_minimal()

glm_plot <- ggarrange(sd, sr, common.legend = TRUE, 
          labels = c("Evenness", "Richness"))

# ggsave("../Wytham_rodent_manuscript/revisions/new_figures/Figure_5_glm_results_2022-06.pdf", glm_plot, width = 8, height = 4)
```

```{r}

######### Plotting predictors

timeseries_df$cal_date <- as.Date(date_decimal(timeseries_df$dec_dates))
timeseries_df$normalised_AS_diversity <- timeseries_df$AS_diversity/max(na.omit(timeseries_df$AS_diversity))
timeseries_df$normalised_MG_diversity <- timeseries_df$MG_diversity/max(na.omit(timeseries_df$MG_diversity))


date_scale_v1 <- scale_x_date(limits = c(ymd("2016/06/31"), ymd("2017/12/12")), date_breaks = "4 months")
date_scale_v2 <- scale_x_date(limits = c(ymd("2017/01/31"), ymd("2017/12/12")))

AS_color = "#66C2A5"
MG_color = "#FC8D62"
MNKA_scale = scale_y_continuous(limits=c(-0.5,28), breaks=seq(0,25,5))
sd_scale = scale_y_continuous(limits=c(-0.05,1.05), breaks=seq(0,1,0.25))
sr_scale = scale_y_continuous(limits=c(1,4.2), breaks=seq(1.0,4.0,1.0))


temp.p <- 
  timeseries_df %>%
  ggplot(aes(cal_date, temp_mean)) +
  geom_point() + geom_line() +
  theme_bw() + 
  xlab("Time") + ylab("Temperature") + date_scale_v1

rain.p <-
  timeseries_df %>%
  ggplot(aes(cal_date, rain_wk)) +
  geom_point() + geom_line() +
  theme_bw() + 
  xlab("Time") + ylab("Rain") + date_scale_v1

humidity.p <- 
  timeseries_df %>%
  ggplot(aes(cal_date, hum)) +
  geom_point() + geom_line() +
  theme_bw() + 
  xlab("Time") + ylab("Humidity") + date_scale_v1

AS.density <- 
  timeseries_df %>%
  ggplot(aes(cal_date, AS_density)) +
  geom_point(color = AS_color) + geom_line(color = AS_color) +
  theme_bw() + 
  xlab("Time") + ylab("MNKA per ha") + date_scale_v2 + MNKA_scale

MG.density <- 
  timeseries_df %>%
  ggplot(aes(cal_date, MG_density)) + 
  geom_point(color = MG_color) + geom_line(color = MG_color) +
  theme_bw() + 
  xlab("Time") + ylab("MNKA per ha") + date_scale_v2 + MNKA_scale

AS_sd <- 
  timeseries_df %>%
  ggplot(aes(cal_date, normalised_AS_diversity)) + 
  geom_point(color = AS_color) + geom_line(color = AS_color) +
  theme_bw() + 
  xlab("Time") + ylab("Normalised Shannon diversity") + date_scale_v2 + sd_scale


AS_sr <- 
  timeseries_df %>%
  ggplot(aes(cal_date, AS_rich)) + 
  geom_point(color = AS_color) + geom_line(color = AS_color) +
  theme_bw() + 
  xlab("Time") + ylab("Species richness") + date_scale_v2 + sr_scale

MG_sd <- 
  timeseries_df %>%
  ggplot(aes(cal_date, normalised_MG_diversity)) + 
  geom_point(color = MG_color) + geom_line(color = MG_color) +
  theme_bw() + 
  xlab("Time") + ylab("Normalised Shannon diversity") + date_scale_v2 + sd_scale

MG_sr <- 
  timeseries_df %>%
  ggplot(aes(cal_date, MG_rich)) + 
  geom_point(color = MG_color) + geom_line(color = MG_color) +
  theme_bw() + 
  xlab("Time") + ylab("Species richness") + date_scale_v2 + sr_scale


(temp.p | rain.p | humidity.p) / (AS.density + AS_sd + AS_sr) / (MG.density + MG_sd + MG_sr)  

library(lubridate)

rain.p <-
  weather_weekly_combined %>%
  ggplot(aes(min_date, rain_cts_cum_week*0.518)) +
  geom_point(pch = 1, colour="black")+
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cc"), se = FALSE, colour="black")+
  theme_bw() +
  scale_x_date(limits=c(dmy("01/01/2017"), dmy("31/12/2017")), breaks=c(dmy("01/04/2017"), dmy("01/07/2017"), dmy("01/10/2017")), date_labels = "%b")+
  scale_y_continuous(expand = c(0,0),lim = c(0,60), breaks = seq(0,70,by=10))+
  labs(x = 'date', y = 'weekly cumulative rainfall (mm)')+
  scale_color_manual(values = c('black','grey','grey'), 
                     labels = c('daily mean','min', 'max'))


weekly_temp = weather_weekly_combined[,c('week','min_date','temp_mean_daily','temp_min_daily', 'temp_max_daily')]
wk_temp_long <- melt(weekly_temp,
                     id.vars=c("week",'min_date'),
                     variable.name="weather_measurement",
                     value.name="value")

temp.p <-
  ggplot(wk_temp_long, aes(min_date, value, color= weather_measurement))+
  geom_point(pch = 1)+
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cc"), se = FALSE)+
  theme_bw()+
  scale_y_continuous(expand = c(0,0),lim = c(-5,25), breaks = seq(-5,25,by=5))+
  #scale_y_continuous(expand = c(0,0),lim = c(-5,25), breaks = seq(-5,25,by=5))+
  labs(x = 'date', y = 'temperature (C)')+
  scale_x_date(limits=c(dmy("01/01/2017"), dmy("31/12/2017")), breaks=c(dmy("01/04/2017"), dmy("01/07/2017"), dmy("01/10/2017")), date_labels = "%b")+
  scale_color_manual(values = c('black','grey','grey')) + theme(legend.position = "none")


weekly_hum = weather_weekly_combined[,c('week','min_date','hum_mean_daily','hum_min_daily', 'hum_max_daily')]
wk_hum_long <- melt(weekly_hum,
                    id.vars=c("week",'min_date'),
                    variable.name="weather_measurement",
                    value.name="value")

hum.p <-
  ggplot(wk_hum_long, aes(min_date, value, color= weather_measurement))+
  geom_point(pch = 1)+
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cc"), se = FALSE)+
  theme_bw()+
  scale_x_date(limits=c(dmy("01/01/2017"), dmy("31/12/2017")), breaks=c(dmy("01/04/2017"), dmy("01/07/2017"), dmy("01/10/2017")), date_labels = "%b")+
  labs(x = 'date', y = 'humidity (%)')+
  scale_color_manual(values = c('black','grey','grey'), 
                     labels = c('daily mean','min', 'max')) + theme(legend.position = "none")

library(patchwork)

temp.p|hum.p|rain.p


test.s <-
all_virus_species.df %>%
  group_by(Species, Interval, virus_species) %>%
  summarise(v_count = sum(viral_read_count)) %>%
  filter(v_count > 0) %>%
    group_by(Species, Interval) %>%
    summarise(n_genera = n())

length(unique(all_virus_species.df$virus_species))


  select_phage <- 
  get_relAbundance_s(unique(c(phage_species.df$virus_species)), data) %>%
  select(virus_species, type, Species, Interval, viral_read_count) %>% 
  group_by(virus_species, type, Species, Interval) %>%
  summarise(sum = sum(viral_read_count)) %>%
  filter(sum >= 1) %>%
  filter(virus_species != "environmental samples") %>%
  group_by(virus_species) %>%
  summarise(sum = sum(sum)) %>%
  filter(sum >= 40)
  
    select_vert <- 

  get_relAbundance_s(unique(c(verterb_virus_species.df$virus_species)), data) %>%
  select(virus_species, type, Species, Interval, viral_read_count) %>% 
  group_by(virus_species, type, Species, Interval) %>%
  summarise(sum = sum(viral_read_count)) %>%
  filter(sum >= 1) %>%
  filter(virus_species != "environmental samples") %>%
   group_by(virus_species) %>%
  summarise(sum = sum(sum)) %>%
  filter(sum >= 10)
  
    test.s <-
      get_relAbundance_s(
        unique(c(select_phage$virus_species, select_vert$virus_species)), data) %>%
      select(virus_species, type, Species, Interval, viral_read_count) %>%
      group_by(virus_species, type, Species, Interval) %>%
  summarise(sum = sum(viral_read_count)) %>%
  filter(sum >= 1) %>%
  filter(virus_species != "environmental samples")   
    
    length(unique(select_phage$virus_species))
    
# write_csv2(test.s, "../Wytham_rodent_manuscript/revisions/summary_table_virus_genera_by_read_count_figure_3.csv")

```


```{r}

host_pop_density.f <-
MNKA %>%
  select(Collection_date, MNKA_per_ha, Species) %>%
  rename(Collection.date = Collection_date) %>%
  mutate(Species = fct_relevel(Species, c("AS", "AF", "MG"))) %>%
  ggplot(aes(as.Date(Collection.date, format = '%d/%m/%Y'), MNKA_per_ha, group = Species, colour = Species)) +
  geom_point() + facet_grid(~Species) + geom_line() + xlab("Time") + ylab("MNKA per ha") +
  scale_colour_manual(values=c("#66C2A5", "#8DA0CB", "#FC8D62")) + theme_minimal() + theme(legend.position = "none") +   scale_x_date(limits =c(dmy("01/11/2016"), dmy("15/12/2017")), breaks=c(dmy("01/01/2017"), dmy("01/04/2017"), dmy("01/07/2017"), dmy("01/10/2017")), date_labels = "%b")


# ggsave(plot = host_pop_density.f, filename =  "../Wytham_rodent_manuscript/revisions/new_figures/Figure_SX_host_pop_density.pdf", width = 8, height = 4)
#   

```
